---
title: "DS250 - Introduction to Data Science"
author: "Student Name"
date: "November 15, 2015"
output: pdf_document
---
```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
```
```{r}
rm( list = ls())
```

# Assignment 7 - Data Visualization
This assignment covers principles of data visualization, concepts of perception, and creating visualizations using R.

### The objectives of this assignment are:
* Review and analysis of visualizations
* Visual perception exercises
* Visualization evaluation
* Creating plots using R

***
#### Question 1 - Visual Analysis
Briefly describe some of the core components of visual analysis:

a) Make it easy to read.

b) Use multiple senses.

c) Understanding the weight of different visuals. I.E. Positioning importance vs. color importance.

***
#### Question 2 - Visual Analysis
What is the 'color wheel' and how is it used?

````
the color wheel makes it easy to distinguish colors that are complimentary and contradictory colors.
````

***
#### Question 3 - Customizing Plots in R
**Using R**, create a Dendrogram plot that matches the following with the sample data:

```{r, echo=TRUE}
data = data.frame(USArrests[,1:3])
dt = dist(data, method = "euclidean")
clust = hclust(dt)
# very simple dendrogram
#plot(clust)
```

a) Modify the title: "US Arrests by State"  
plot(clust, main = "US Arrests by State")
b) Modify the x and y axis labels
plot(clust, main = "US Arrests by State", xlab = "state", ylab="value")

***
#### Question 4 - Visual Evaluation
Examine the infographic:
http://www.informationisbeautifulawards.com/showcase/611-the-graphic-continuum  

Provide feedback on the following:
a) Organization
b) Typography  
c) What 'story' does it tell?  
d) Any suggestions or feedback on improvements?  

````
A: The chart feels a little bit larger than it really needs to be. I feel like a lot of the space of the chart is used for lines that connect a few similar concepts that are only loosely related. Seems like it would be a little easier to read if it wasn't so spaced out. Where the items are on the page is also meaningless.
B: I really love simple font and muted colors. I find it very easy to read and the colors of the different types of visualizations is clear as well.
C: The story is that there are many types of visualizations and they all have different uses.
D: Make the information more compact and the images larger. I find the lines with different information on it superfluous and seeing the visualizations as larger images would convey the information better to me. 
````

***
#### Question 5 - Visual Evaluation
Examine the following infographic:
http://www.informationisbeautifulawards.com/showcase/490-world-cup-2014-wall-chart

Provide feedback on the following:
a) Organization  
b) Typography  
c) What 'story' does it tell?  
d) Any suggestions or thoughts on improvements?

````
A: Initially I was a kind of overwhelmed by the bracket information being provided in a circle but once I got over it I really liked seeing the groups and being able to see the group standings underneath, I thought that was a really elegant way to get both of those pieces of information on the same chart.
B: I really dislike upsidedown text, no one can read that! Everything else about the typography is easy and straightforward.
C: I like that all the information feeds into the middle where the World Cup Winner is eventully put. Guiding the reader in from all around the circle to the middle is really clean and clear. 
D: No upsidedown text.
````

***
#### Question 6 - Histograms and Density Plots
Review and update the following R code to produce the following using qplots:  
a) Histogram  
b) Density plot  

```{r, echo=TRUE}
library(ggplot2)  
#qplot(Petal.Length, data=iris, geom="")  ## Histogram  
qplot(Petal.Length, data=iris, geom="histogram", binwidth=.5)
#qplot(Petal.Length, data=iris, geom="")  ## Density plot 
qplot(Petal.Length, data=iris, geom="density")
```

***
#### Question 7 - Create Bar Chart
From the dataset below, create a **bar chart** using barplot:

```{r, echo=TRUE}
ds <- c(2,3,3,3,4,5,5,5,5,6)
counts <- table(ds)
counts
barplot(counts)
```

a) Add a title and x and y axis labels  
barplot(counts, xlab="How Good I felt At the End of the Day", ylab="Candy Bars Eaten that Day")
b) Change the color of the bar plot bars
barplot(counts, xlab="How Good I felt At the End of the Day", ylab="Candy Bars Eaten that Day", col=c("blue", "red", "green"))
c) Use the 'density' method to alter the bar plot bars  
barplot(counts, xlab="How Good I felt At the End of the Day", ylab="Candy Bars Eaten that Day", col=c("blue", "red", "green"), density = 10)
An award winning graph if you were to ask me.
***
#### Question 8 - Visualizations in R
From the dataset below, create a **line chart** showing (CO2 emissions per capita) by (year):

```{r, echo=TRUE}
year <- c("1961","1962","1963","1964","1965","1966","1967","1968","1969","1970","1971","1972","1973","1974","1975","1976","1977","1978","1979","1980","1981","1982","1983","1984","1985","1986","1987","1988","1989","1990","1991","1992","1993","1994","1995","1996","1997","1998","1999","2000","2001","2002","2003","2004","2005","2006","2007","2008","2009","2010")

co2_emission_per_cap_qt <- as.numeric(c("0.836046900792028","0.661428164381093","0.640001899360285","0.625646053941047","0.665524211218076","0.710891381561055","0.574162146975018","0.60545199674633","0.725149509123457","0.942934534989582","1.04223969658961","1.08067663654397","1.09819569131687","1.09736711056811","1.25012409495905","1.28528313553995","1.38884289658754","1.52920110964112","1.5426750986837","1.49525074931082","1.46043181655825","1.56673968353113","1.62905590778943","1.75044806018373","1.87105479144466","1.93943425697654","2.03841068876927","2.1509052249848","2.15307791087471","2.16770307659104","2.24590127565651","2.31420729031649","2.44280065934625","2.56599389177193","2.75575496636525","2.84430958153669","2.82056789057578","2.67674598026467","2.64864924664833","2.69686243322549","2.74212081298895","2.88522504139331","3.51224542766222","4.08013890554173","4.4411506949345","4.89272709798477","5.15356401658718","5.31115185538876","5.77814318390097","6.19485757472686"))

df.china_co2 <- data.frame(year,co2_emission_per_cap_qt)
head(df.china_co2,5)
ggplot(df.china_co2, aes(year, co2_emission_per_cap_qt, group=1)) + geom_line()
```

***
#### Question 9 - Scatter Plots
Review and modify the following scatter plot:

```{r, echo=TRUE}
require(ggplot2)
qplot(dose, len, data=ToothGrowth, geom="point")
```

a) display the (supp) values (OJ,VC) points with different colors  
qplot(dose, len, data=ToothGrowth, color=supp, geom="point")
b) separate the (supp) values (OJ,VC) into separate plots on the same chart  
qplot(dose, len, data=ToothGrowth, color=supp, geom="point") + geom_point(position=position_dodge(width=.2))
c) add a trend line using (geom=c("point","smooth"))  
qplot(dose, len, data=ToothGrowth, color=supp, geom=c("point","smooth"))

***
#### Question 10 - Bubble Plots
Using the data set below, create a bubble plot showing x and y and variable sizes

```{r, echo=TRUE}
set.seed(57)        

x_var <- rnorm( n = 15, mean = 5, sd = 2)
y_var <- x_var + rnorm(n = 15, mean = 5, sd =4)
size_var <- runif(15, 1,10)

df.test_data <- data.frame(x_var, y_var, size_var)
head(df.test_data,5)
qplot(x_var, y_var, data=df.test_data, size=size_var)
```

#### **END**
