---
title: "DS250 - Introduction to Data Science"
author: "Sam Robinson"
date: "October 04, 2016"
output: html_document
---
# Assignment 1
This assignment covers some beginning analysis exercises and an introduction to R programming that covers basic operations to get you started in R scripting.

### The objectives of this assignment are:
* Review of data science project lifecycle
* Review of data science project roles
* Simple analysis excercises
* Experience in the R/R-studio development environment
* Practice using R vectors and vector operations
* Crash course in Markdown document formatting

***
#### Question 1 - Project Lifecycle
1. Beneath each item below, write a short description for each of the following project activities:
+ Problem Statement
```{r eval=FALSE}
- A brief description of the thing you want to test and solve.

````
+ Data Profile & Discovery
```{r eval=FALSE}
- Collecting the data you believe is needed for your problem statement (and then collecting it again when you realize you may be missing the mark)
````
+ Hypothesis Statement
```{r eval=FALSE}
-The way in which you think you can solve the stated problem.
````
+ Data Preparation
```{r eval=FALSE}
- This is also cleaning the data. Refactoring or reorganizing the data to help put it into a state that can be analyzed and removing any bad data that might creep into your data set.
````
+ Model Development
```{r eval=FALSE}
- An expression of how the data you are collecting will help answer the question being asked. I.E. Players in a game (real life example) who do a certain action during the first time they play the game have a higher chance of playing the game for longer.
````
+ Model Evaluation
```{r eval=FALSE}
- Does the model and data collected support your hypothesis for the problem statement? Is there another way to prove out your hypothesis?
````

***
#### Question 2 - Project Roles
1. Breifly describe the following **roles** in a Data Science project:
+ Product Manager
```{r eval=FALSE}
- Evaluates costs and timelines for your data science project. Helps report to outside entities and insolates the team from distractions.
```
+ Domain Expert
```{r eval=FALSE}
- The person who has the most knowledge about the space your data is being collected from. Like an expert on farming or the lead designer of a video game.
```
+ Data Engineer
```{r eval=FALSE}
- This is the person maintaing the data base and data collection. 
  ```
+ Data Scientist
```{r eval=FALSE}
- Help create predictive models and testable hypothesis for the stated problem you are trying to solve.
```
+ Graphic Designer
```{r eval=FALSE}
- The person helping create a visualization and/or presentation for the results of the Data Science Project.
```

***
#### Question 3 - Comparision: Business Intelligence & Data Science
1. List some common objectives and activities in Business Intelligence:
+ Return on investement
+ Longer term customer value.
+ Server and developer costs.


2. List some objectives and activities in Data Science:
+ Create testable hypothesis about a product.
+ Attempt to find correlation and causation of app user actions to long term value. (Users who hit button X and up seeing more adds than users who hit button Y)
+ Create predicitive models.

***
#### Question 4 - Simple Data Analysis - World Population Demographics
Using the [World Bank - Population Data Site](http://data.worldbank.org/indicator/SP.POP.TOTL), complete the following analysis tasks: *(You may use any tools for this excercise)*

+ Identify the top three countries with the **largest population growth** in the past five years.
```{r eval=FALSE}
Enter list of three countries and population counts here...
Oman
Qatar
Lebanon
```
+ Identify the top three countries with the **lowest (or negative) population growth** in the past five years.
```{r eval=FALSE}
Enter list of three countries and population counts here...
Andorra - -3.510073239
Syrian Arab Republic - -2.007556716
Lithuania - -1.598000073
```

***
#### Question 5 - R Scripting: Using Vectors
1. For the following vectors:
```{r eval=FALSE}
a <- c(2,3,4,5,6,7)
b <- c(1,2)
```

Find the value for the vector:
```{r eval=FALSE}
d <- a*b
d = c(2,6,4,10,6,14)
```

Results:
```{r eval=FALSE}
Your results here:??? I don't have a question on my sheet here.
  
```

***
#### Question 6 - R Scripting: Using Functions
1. For the following functions:

```{r eval=FALSE}
y <- 3
f <- function(x) { y <- 2 
                   y ^  2 + g(x)}
g <- function(x) { x * y }
```

find the value of **f(6)**:

Results:
```{r eval=FALSE}
Your results here:
f(6) = 22
```

***
#### Question 7 - R Scripting: Using Apply
1. Use the lapply() function to compute the mean for the following vectors combined together:
```{r eval=FALSE}
a <- c(1,1,1,1,2,2,2,2,2)
b <- c(10,12,15,12,NA,30,42,38,40)
```

Results:
```{r eval=FALSE}
Your results here:
```

2. Write a statement to remove the **'NA'** from the (b) vector above.

Results:
```{r eval=FALSE}
Your results here:
fx <- function(x) {x <- na.omit(x)}

```

***
#### Question 8 - R Scripting: Vector math functions
1. Given the data below, write the command to complete the following operations:
```{r eval=FALSE}
x = c(1,3,5,7,9)
y = c(2,3,5,7,11,13)
```
+  x+1
answer <- x+1
+  y*2
answer <- y*2
+  length(x) and length(y)
answerInteger <- length(x)
answerInteger <- length(y)
+  x + y
xy <- x+y
Gives a warning
Warning message:
In x + y : longer object length is not a multiple of shorter object length

+  sum(x>5) and sum(x[x>5])

sum(x>5) = 2L
sum(x[x>5]) = 16

Here is a function that I wrote that I didn't need to write, but it was good practice.
sumVal <- function(x) {
    retVal <- 0
    for(i in x){
        if(i > 5){
            retVal <- retVal + i
        }
    }
    return(retVal)
}

sumVal(x)
sumVal(x[x>5])


+  sum(x>5 | x< 3)
+  y[3]
+  y[-3]
+  y[x]
+  y[y>=7]

***
#### Question 9 - Collaboration - References & Resources Forum Posting
1. Research and post an interesting article on a data science related topic to the [References & Resources](https://canvas.uw.edu/courses/1092104/discussion_topics/3415427) forum. Include the following details:
+ Why you chose this topic
+ Briefly list the key points of the article
+ URL to the article isn the post

***
#### Question 10 (Optional) - Probability Puzzle
1. **Single Dice Game** *(You may use any tools for this excercise)*
+ You are playing a game with one (1) standard six-sided dice
+ You may roll the dice three (3) times
+ You will receive a reward of the amount you roll: (1-6)
+ You cannot keep prior rolls

2. What is the expected average payout you will receive playing this game?
+ (Describe your answer)

The expected value of any single dice roll is equal to mean of numbers 1 through 6, assuming we are using a fair die. That equals 3.5. The expected value of each dice roll does not change based on the previous dice roll, therefore as a result the expected dice roll is 3*(mean(1:6)) or 10.5.

***

##### **END**