---
title: "Intro to Machine Learning"
date: "October 25, 2015"
output:  html_document
---
```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
```
```{r}
rm( list = ls())  # Clear environment
```

# Lab 6.1 - Linear Regression

#### **Univariant Regression Analysis**
#### **Multivariant Regression Analysis**


## 1\. Regression Review

**Anscombe** sample data set:

````{r}
data(anscombe)
attach(anscombe)
head(anscombe,5)
````

a\. Correlation of x1 and y1: `r cor(x1, y1)`  
b\. Correlation of x2 and y2: `r cor(x1, y2)` 

#### Plot of variables -- Look for outliers

````{r}
par(mfrow=c(2,2)) #create a 2x2 grid for plotting
plot(x1, y1, main="Plot 1")
plot(x2, y2, main="Plot 2")
plot(x3, y3, main="Plot 3")
plot(x4, y4, main="Plot 4")
````

***
## 2\. Simple Linear regression

#### Load **Snake River** data set:

````{r}
#install.packages("alr3")
library(alr3)
data(snake)
names(snake) = c("content", "yield")
attach(snake)
head(snake,5)
plot(content, yield, xlab="water content of snow", ylab="water yield")
````

#### Create linear model

````{r}
yield.fit = lm(yield~content)
summary(yield.fit)
````

a\. **Yield** is **0.72538** plus **0.49808** times the content.  
b\. **p-value** = (**4.632e-08**) which is much less than 0.05 so H0 is significant

#### Plot Regression

````{r}
plot(content, yield)
abline(yield.fit, lwd=3, col="red")
````

#### Plot Fit Analysis

````{r}
par(mfrow=c(2,2))
plot(yield.fit)
````


#### Plot Fit Confidence Intervals

````{r}
qqPlot(yield.fit)
````

***
## 3\. Multivariate linear regression

#### Load sample data set (water)

````{r}
data(water)
socal.water = water[ ,-1] # Exclude Year
head(socal.water,5)
````

#### Exploratory analysis of correlated variables

````{r}
# install.packages("corrplot")
library(corrplot)
water.cor = cor(socal.water)
corrplot(water.cor, method="ellipse")
````

#### Feature Selection

a\. We will first build and examine a model with all the features, then drill down with best subsets to select the best fit.

````{r}
# install.packages("leaps")
library(leaps)
fit=lm(BSAAM~., data=socal.water)
summary(fit)
````

b\. Create fit subsets:

````{r}
sub.fit = regsubsets(BSAAM~., data=socal.water)
best.summary = summary(sub.fit)
names(best.summary)
````

c\. Find minimum features:

````{r}
which.min(best.summary$rss)
````

d\. Compare fit results using Mallow's Cp:

````{r}
par(mfrow=c(1,2))
plot(best.summary$cp, xlab="number of features", ylab="cp")
plot(sub.fit, scale="Cp")
````

e\. Create new model on specific features:

````{r}
best.fit = lm(BSAAM~APSLAKE+OPRC+OPSLAKE, data=socal.water)
summary(best.fit)
````

With the three-feature model, F-statistic and all the t-tests have significant p-values. 

#### Plot model fit analysis
````{r}
par(mfrow=c(2,2))
plot(best.fit)
````

f\. Investigate colinearity

````{r}
vif(best.fit)
````

From the adjusted R-Squared values, the two-variable model of **APSLAKE** and **OPSLAKE** produced a value of 0.90, while adding **OPRC** only marginally increased it to 0.92:

````{r}
 plot(socal.water$OPRC, socal.water$OPSLAKE, xlab="OPRC", ylab="OPSLAKE")
````

g\. Drop the **OPRC** feature and compare predicted vs actual

````{r}
fit.2 = lm(BSAAM~APSLAKE+OPSLAKE, data=socal.water)

plot(fit.2$fitted.values, socal.water$BSAAM,xlab="predicted", ylab="actual", main="Predicted vs.Actual")
````

#### Plot presentation

````{r}
socal.water["Actual"] = water$BSAAM #create the vector Actual
socal.water["Forecast"] = NA #create a vector for the predictions named Forecast
socal.water$Forecast = predict(fit.2) #populate Forecast with the predicted values

library(ggplot2)
ggplot(socal.water, aes(x=Forecast, y=Actual)) +geom_point() + geom_smooth(method=lm) + labs(title = "Forecast versus Actuals")
````


#### **End**
